---
title: "Empirical Analysis"
author: FIRE199 Sustainability Analytics
format: html
editor: visual
---

## Chen & Whalley (2012) Basic OLS Model

$log(AirQuality_t) = \gamma_0 + \gamma_1 MetroRidership_t + \gamma_2 x_t + e_t$

$t$ is time

Controls $x_t$ include:

-   gas content regulation dummy

-   $temp$, $temp_{-1}$, $temp_{-1}^2$, $temp_{-1}^3$, $temp_{-1}^4$ and same thing for wind speed

## Chen & Whalley (2012) DB-OLS Model

$log(AirQuality_t) = \delta_0 + \delta_1 MetroOpen_t + \delta_2 x_t + \delta_3 P(t) + \delta_4 P(t) \times MetroOpen_t + e_t$

Controls $x_t$ include:

-   gas content regulation dummy

-   $temp$, $temp_{-1}$, $temp_{-1}^2$, $temp_{-1}^3$, $temp_{-1}^4$ and same thing for wind speed and humidity

-   month FE

-   day of week FE

-   hour FE

-   hour FE $\times$ day of week FE

$P(t)$ is $t$, $t^2$, $t^3$

Make a table of average pollution before and after

```{r}
library("tidyverse")

setwd("C:/Git Local/research-light-rail")

df<-read.csv("pm25_highway_daily.csv") %>%
  filter(Address=="Charlotte, NC")
```

Make date column

```{r}

```

Make date variables

```{r}

```

Make table summarizing pm2.5 before and after

```{r}

```

```{r}

```

## Xie et al. (2019)

Also controlled for:

-   holiday status
