---
title: "Do Light Rails Provide the Track to Cleaner Air?"
date: 9/10/2024
author: FIRE Sustainability Analytics
format: gfm
editor: visual
---

# Introduction

This research project aims to find if light rail openings result in an overall decrease in air pollution in U.S. urban areas? The answer to this question will help predict air pollution impacts of the Purple Line in Maryland that is scheduled to open in 2027.

To examine the potential impact of light rail openings, we will analyze the outcomes in other U.S. cities with light rail systems, utilizing ground-level PM2.5 data developed by Di et al. (2021). This data has a spatial resolution of 1-kilometer grids and a daily temporal resolution from 2000 to 2016. We will focus on cities where light rail is the primary mode of rail transit, and no other urban rail systems are present. Four light rail systems in the United States meet these criteria: the Lynx Blue Line in Charlotte, North Carolina; the Red METRORail Line in Houston, Texas; the Blue METRO Line in Minneapolis-St. Paul, Minnesota; and the Valley Metro Rail in Phoenix-Mesa, Arizona.

We will employ a difference-in-difference estimation strategy to compare air pollution trends in cities with light rail openings and control cities that had similar air pollution trends before the light rail openings. This will involve determining monthly average PM2.5 levels for all urban areas in the U.S. We will then use synthetic control methods to generate data for control cities with pre-light rail pollution trends similar to each treated city. Subsequently, we will calculate potential changes in air pollution due to light rail openings.

This research project aims to provide the most rigorous impact estimation of light rail openings on air pollution to date. While existing studies suggest that light rails are environmentally superior to other transportation options, they haven't conclusively shown that their opening reduces city-wide air pollution (Mulley et al., 2017; Ham et al., 2017).

Unlike numerous studies on subway openings, there are limited studies on light rails. Park & Sener (2019) indicate a 24% decrease in CO measured from monitors after the opening of the Houston light rail, and Fageda (2021) suggests a 3% decrease in PM2.5 after light-rail openings in European cities.

Our study will differ from existing studies on light rail and air pollution in several ways. Because Park & Sener (2019) analyzed the change in pollution in only one city, there may be other factors affecting the change in the pollution that confounds with the opening of the light rail. Our study will also use higher frequency pollution data than the annual level used in Fageda (2021), allowing us to control for meteorological factors. 

As Park & Sener (2019) only included data two years before the light rail opening, and Fageda (2021) included data three years before the light rail openings, their findings can confound with the construction period of light rails. As shown in Figure 1, light rail construction usually happens around three years before light rail openings. There is also an increase in PM2.5 levels during the construction period. By including data only two to three years before light rail openings, existing studies can overestimate the degree of pollution reduction, as construction activities can contribute to the higher pollution levels before light rail openings. 

## Treatment Criteria

-   As Di et al. (2021) provides data from 2000 to 2016, candidate treatment cities must have light rails that opened from January 2004 to December 2012 in order to have four years of data before and after light rail opening.

-   Treatment cities must have no increase heavy rail usage in the four years before and after opening window.

+-----------------+---------------+------------------------+------------+-----------+
| City            | Opening Month | Construction           | Eval Start | Eval Ends |
+=================+===============+========================+============+===========+
| Charlotte, NC   | Nov 2007      | Mar 2005 - Nov 2007,   | Nov 2003   | Nov 2011  |
|                 |               |                        |            |           |
|                 |               | Dec 2012 onwards       |            |           |
+-----------------+---------------+------------------------+------------+-----------+
| Houston, TX     | Jan 2004      | March 2001 - Jan 2004, | Jan 2000   | Jan 2008  |
|                 |               |                        |            |           |
|                 |               | Jan 2014 onwards       |            |           |
+-----------------+---------------+------------------------+------------+-----------+
| Twin Cities, MN | Jun 2004      | Feb 2001 - Jun 2004    | Jun 2000   | Jun 2008  |
|                 |               |                        |            |           |
|                 |               | Jan 2010 onwards       |            |           |
+-----------------+---------------+------------------------+------------+-----------+
| Phoenix, AZ     | Dec 2008      | July 2005 - Dec 2008   | Dec 2004   | Dec 2012  |
|                 |               |                        |            |           |
|                 |               | July 2012 onwards      |            |           |
+-----------------+---------------+------------------------+------------+-----------+

## Control cities

The plot below shows number of trips for each mode of revenue-generating transportation in Charlotte.

Because buses are the primary alternative transportation in the city, control cities for **Charlotte** are cities that primarily runs on buses in South Carolina and North Carolina with no rails and no light rails.

```{r}
#| echo: false
#| message: false
#| warning: false

library("tidyverse")

trans<-read.csv("ntd_04to13.csv") 

charlotte<-trans |>
  filter(ID==4008) 

ggplot(data=charlotte)+
  geom_line(aes(y=LR, x=year, col="light rail"))+
  geom_line(aes(y=bus, x=year, col="bus")) + 
  geom_line(aes(y=cars, x=year, col="cars")) + 
  geom_line(aes(y=rail, x=year, col="rail")) + 
  geom_line(aes(y=oth, x=year, col="other")) + 
  ylab("Number of Trips for Each Public Transport Mode")+
  ggtitle("Charlotte, NC")+
  geom_vline(xintercept=2007, linetype = "dotted")+
  theme_bw()
```

The plot below shows number of trips for each mode of revenue-generating transportation in Houston.

Control cities for **Houston** are other cities in Texas with no rails, no light rails, and less than 5% of other forms of other transportation in light rail opening year.

```{r}
#| echo: false
#| message: false
#| warning: false

trans<-read.csv("ntd_04to13.csv") 

houston<-trans |>
  filter(ID==6008) 

ggplot(data=houston)+
  geom_line(aes(y=LR, x=year, col="light rail"))+
  geom_line(aes(y=bus, x=year, col="bus")) + 
  geom_line(aes(y=cars, x=year, col="cars")) + 
  geom_line(aes(y=rail, x=year, col="rail")) + 
  geom_line(aes(y=oth, x=year, col="other")) + 
  ylab("Number of Trips for Each Public Transport Mode")+
  ggtitle("Houston, TX")+
  geom_vline(xintercept=2004, linetype = "dotted")+
  theme_bw()
```

The plot below shows number of trips for each mode of revenue-generating transportation in the Twin Cities

Control cities for the **Twin Cities** are other cities in Minnesota and Wisconsin with no rails, no light rails, and less than 5% of other forms of other transportation in light rail opening year.

```{r}
#| echo: false
#| message: false
#| warning: false

trans<-read.csv("ntd_04to13.csv") 

mn<-trans |>
  filter(ID==5027) 

ggplot(data=mn)+
 geom_line(aes(y=LR, x=year, col="light rail"))+
  geom_line(aes(y=bus, x=year, col="bus")) + 
  geom_line(aes(y=cars, x=year, col="cars")) + 
  geom_line(aes(y=rail, x=year, col="rail")) + 
  geom_line(aes(y=oth, x=year, col="other")) + 
  ylab("Number of Trips for Each Public Transport Mode")+
  geom_vline(xintercept=2004, linetype = "dotted")+
  ggtitle("Twin Cities, TX")+
  theme_bw()
```

The plot below shows number of trips for each mode of revenue-generating transportation in the Phoenix-Mesa.

Control cities for **Phoenix-Mesa** are cities in other Arizona metropolitan areas with no rails and no light rails.

```{r}
#| echo: false
#| message: false
#| warning: false

id_az<-c(9032, 9034, 9131, 9136, 9172, 9169, 9132, 9140, 9209)

phoenix<-trans |>
  filter(ID %in% id_az) |>
  group_by(year) |>
  summarize(bus=sum(bus), 
            cars=sum(cars), 
            LR=sum(LR), 
            rail=sum(rail), 
            oth=sum(oth),
            total=sum(total)) |>
  mutate(share_bus=bus/total, share_rail=rail/total, share_light=LR/total, 
         share_oth=oth/total, share_cars=cars/total)

ggplot(data=phoenix)+
  geom_line(aes(y=LR, x=year, col="light rail"))+
  geom_line(aes(y=bus, x=year, col="bus")) + 
  geom_line(aes(y=cars, x=year, col="cars")) + 
  geom_line(aes(y=rail, x=year, col="rail")) + 
  geom_line(aes(y=oth, x=year, col="other")) + 
  ylab("Number of Trips for Each Public Transport Mode")+
  ggtitle("Phoenix-Mesa, AZ")+
  geom_vline(xintercept=2008, linetype = "dotted")+
  theme_bw()
```

## Data

### PM2.5

As particulate matter is one of the most damaging air pollutants, we use ground-level PM2.5 concentrations to represent city-level air quality. PM2.5 data comes from Di et al. (2019), which provides daily PM2.5 concentrations in grid cells at a resolution of 1 km for the years 2000 to 2016. We then find the monthly average PM2.5 concentration within a 10 km buffer area around each city's population centroid for the three treatment cities: Charlotte, Houston, Texas, and Minneapolis-St. Paul. As the Valley Metro Rail connects two cities, Phoenix and Mesa, Arizona, we find the average PM2.5 concentration for the 10 km buffer areas around both cities' population centroid for each month.

The 10 km radius around the population centroid is the black line, and the 10 km buffer radius around the light rail centroid is the red line.

```{r}
#| message: false

library("maptiles")
library("terra")

#city centroid
cities<-read.csv("allcities_latlon.csv")

df<-cities |>
  filter(Address=="Charlotte, NC") |>
  select(lon, lat)

#converts df into a spatvector
x <- vect(df, geom=c("lon", "lat"), crs="+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ")

#create a 10 km (10,000 meter) buffer (radius)
pts_buffer<-buffer(x, width = 10000)
  
#light rail centroid
char_lr <- vect("G:/Shared drives/2024 FIRE Light Rail/DATA/LYNX_Blue_Line_Route/LYNX_Blue_Line_Route.shp")

#combines entire route into one line
char_lr<-aggregate(char_lr, dissolve=TRUE)

#change coordinate system to match points
lr_project<-project(char_lr, "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ")

#find centroid of light rail route
lrc<-centroids(lr_project, inside=FALSE)

#create 10 km buffer around light rail centroid
pts_buffer1<-buffer(lrc, width = 10000)

#approximate size of the background
extent<-buffer(x, width = 10000)

bg <- get_tiles(ext(extent))

plot(bg)
points(x)
lines(lr_project, col="blue")
lines(pts_buffer1, col="red")
lines(pts_buffer)
```

```{r}
#| echo: false
#| message: false
#| warning: false

df<-read.csv("pm25_allcities.csv") |>
  filter(add=="Charlotte, NC") |>
  mutate(date=as.Date(date))

ggplot(data=df, aes(x=date, y=pm25))+geom_line()+
  ggtitle("Charlotte, NC")+
  ylab("PM2.5 (in ug/m3)")+
  xlab("Year")+
  geom_vline(aes(xintercept=as.Date("2007-11-01"), col="Light Rail Opening"), linetype="dashed")+
  geom_vline(aes(xintercept=as.Date("2003-11-01"), col="Four Years Before or After Opening"), linetype="dotted", size=1)+
  geom_vline(aes(xintercept=as.Date("2011-11-01"), col="Four Years Before or After Opening"), linetype="dotted", size=1)+
  annotate("rect", xmin = as.Date("2005-03-01"), xmax = as.Date("2007-11-01"), 
           ymin = 0, ymax = 25, alpha = .5,fill = "grey")+
  annotate("rect", xmin = as.Date("2012-12-01"), xmax = as.Date("2018-12-01"), 
           ymin = 0, ymax = 25, alpha = .5,fill = "grey")+
  theme_bw()+
  scale_color_manual(values=c("#FFCD00", "#c8102e", "#eeeeee"))

```

```{r}
#| echo: false
#| message: false
#| warning: false

df<-cities |>
  filter(Address=="Houston, TX") |>
  select(lon, lat)

x <- vect(df, geom=c("lon", "lat"), crs="+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ")
pts_buffer<-buffer(x, width = 10000)
  
#light rail centroid
Transit_Map_GEO <- vect("G:/Shared drives/2024 FIRE Light Rail/DATA/National_Transit_Map_Routes/National_Transit_Map_Routes.shp")
trans <- subset(Transit_Map_GEO, Transit_Map_GEO$route_long == "METRORAIL RED LINE")
char_lr<-aggregate(trans, dissolve=TRUE)
lr_project<-project(char_lr, "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ")
lrc<-centroids(lr_project, inside=FALSE)
pts_buffer1<-buffer(lrc, width = 10000)

bg <- get_tiles(ext(pts_buffer))
plot(bg)
lines(lr_project, col="blue")
lines(pts_buffer1, col="red")
lines(pts_buffer)
```

```{r}
#| echo: false
#| message: false
#| warning: false

df<-read.csv("pm25_allcities.csv") |>
  filter(add=="Houston, TX") |>
  mutate(date=as.Date(date))

ggplot(data=df, aes(x=date, y=pm25))+geom_line()+
  ggtitle("Houston, TX")+
  ylab("PM2.5 (in ug/m3)")+
  xlab("Year")+
  geom_vline(aes(xintercept=as.Date("2004-01-01"), col="Light Rail Opening"), linetype="dashed")+
  geom_vline(aes(xintercept=as.Date("2000-01-01"), col="Four Years Before or After Opening"), linetype="dotted", size=1)+
  geom_vline(aes(xintercept=as.Date("2008-01-01"), col="Four Years Before or After Opening"), linetype="dotted", size=1)+
  annotate("rect", xmin = as.Date("2001-03-01"), xmax = as.Date("2004-01-01"), 
           ymin = 0, ymax = 25, alpha = .5,fill = "grey")+
  annotate("rect", xmin = as.Date("2014-01-01"), xmax = as.Date("2018-12-01"), 
           ymin = 0, ymax = 25, alpha = .5,fill = "grey")+
  theme_bw()+
  scale_color_manual(values=c("#FFCD00", "#c8102e", "#eeeeee"))

```

```{r}
#| echo: false
#| message: false
#| warning: false

df<-cities |>
  filter(Address=="Minneapolis-St. Paul, MN") |>
  select(lon, lat)

x <- vect(df, geom=c("lon", "lat"), crs="+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ")
pts_buffer<-buffer(x, width = 10000)
  
#light rail centroid
Transit_Map_GEO <- vect("G:/Shared drives/2024 FIRE Light Rail/DATA/National_Transit_Map_Routes/National_Transit_Map_Routes.shp")
trans <- subset(
  Transit_Map_GEO, 
  Transit_Map_GEO$route_url == "https://www.metrotransit.org/route/blue" | 
    Transit_Map_GEO$route_url == "https://www.metrotransit.org/route/green")
char_lr<-aggregate(trans, dissolve=TRUE)
lr_project<-project(char_lr, "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ")
lrc<-centroids(lr_project, inside=FALSE)
pts_buffer1<-buffer(lrc, width = 10000)

bg <- get_tiles(ext(pts_buffer))
plot(bg)
lines(lr_project, col="blue")
lines(pts_buffer1, col="red")
lines(pts_buffer)
```

```{r}
#| echo: false
#| message: false
#| warning: false

df<-read.csv("pm25_allcities.csv") |>
  filter(add=="Minneapolis-St. Paul, MN") |>
  mutate(date=as.Date(date))

ggplot(data=df, aes(x=date, y=pm25))+geom_line()+
  ggtitle("Minneapolis-St. Paul, MN")+
  ylab("PM2.5 (in ug/m3)")+
  xlab("Year")+
  geom_vline(aes(xintercept=as.Date("2004-07-01"),  col="Light Rail Opening"), linetype="dashed")+
  geom_vline(aes(xintercept=as.Date("2000-07-01"), col="Four Years Before or After Opening"), linetype="dotted", size=1)+
  geom_vline(aes(xintercept=as.Date("2008-07-01"), col="Four Years Before or After Opening"), linetype="dotted", size=1)+
  annotate("rect", xmin = as.Date("2001-02-01"), xmax = as.Date("2004-07-01"), 
           ymin = 0, ymax = 25, alpha = .5,fill = "grey")+
  annotate("rect", xmin = as.Date("2010-01-01"), xmax = as.Date("2018-12-01"), 
           ymin = 0, ymax = 25, alpha = .5,fill = "grey")+
  theme_bw()+
  scale_color_manual(values=c("#FFCD00", "#c8102e", "#eeeeee"))
```

```{r}
#| echo: false
#| message: false
#| warning: false

df<-cities |>
  filter(joint_address=="Phoenix-Mesa, AZ") |>
  select(lon, lat)

x <- vect(df, geom=c("lon", "lat"), crs="+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ")
pts_buffer<-buffer(x, width = 10000)
pmshape<-aggregate(pts_buffer)
  
#light rail centroid
Transit_Map_GEO <- vect(
  "G:/Shared drives/2024 FIRE Light Rail/DATA/Valley_Metro_Light_Rail/LightRailLine.shp")
trans <- subset(Transit_Map_GEO, Transit_Map_GEO$SYMBOLOGY=="METRO")
char_lr<-aggregate(trans, dissolve=TRUE)
lr_project<-project(char_lr, "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ")
lrc<-centroids(lr_project, inside=FALSE)
pts_buffer1<-buffer(lrc, width = 10000)

bg <- get_tiles(ext(pmshape))
plot(bg)
lines(lr_project, col="blue")
lines(pts_buffer1, col="red")
lines(pmshape)
```

```{r}
#| echo: false
#| message: false
#| warning: false

df<-read.csv("pm25_allcities.csv") |>
  filter(add=="Phoenix-Mesa, AZ") |>
  mutate(date=as.Date(date))

ggplot(data=df, aes(x=date, y=pm25))+geom_line()+
  ggtitle("Phoenix, AZ")+
  ylab("PM2.5 (in ug/m3)")+
  xlab("Year")+
  geom_vline(aes(xintercept=as.Date("2008-12-01"), col="Light Rail Opening"), linetype="dashed")+
  geom_vline(aes(xintercept=as.Date("2004-12-01"), col="Three Years Before or After Opening"), linetype="dotted", size=1)+
  geom_vline(aes(xintercept=as.Date("2012-12-01"), col="Three Years Before or After Opening"), linetype="dotted", size=1)+
  annotate("rect", xmin = as.Date("2005-07-01"), xmax = as.Date("2008-12-01"), 
           ymin = 0, ymax = 25, alpha = .5,fill = "grey")+
  annotate("rect", xmin = as.Date("2012-07-01"), xmax = as.Date("2018-12-01"), 
           ymin = 0, ymax = 25, alpha = .5,fill = "grey")+
  theme_bw()+
  scale_color_manual(values=c("#FFCD00", "#c8102e", "#eeeeee"))
```

### Meteorology

To take into account meteorological conditions, we include total precipitation, snowfall, humidity, air density, wind speed, and air temperature from NASA's Modern-Era Retrospective Analysis for Research and Applications version 2 (MERRA-2), which provides these meteorological values every day at a grid cell resolution of 0.625 degrees x 0.5 degrees. We then calculate the monthly average values for each meteorological variable within the 10 km buffer areas of treatment and control cities.

## Synthetic Control

For each city, we minimized the sum of squared residuals for the time period from Jan 2000 to the beginning of light rail construction.

We omitted results from Houston and the Twin Cities because the pre-construction number of observations is very low.

### Charlotte

```{r}
#| echo: false
#| message: false
#| warning: false

library("Synth")
library("gsynth")

p<-read.csv("pm25_allcities.csv")
m<-read.csv("merra_allcities.csv")

df<-merge(p, m, by=c("date", "add", "month", "year"))

cgroup<-c("Charlotte, NC", "Asheville, NC", "Charleston-North Charleston, SC", 
          "Columbia, SC", "Concord, NC", "Durham, NC", "Fayetteville, NC", 
          "Greenville, SC", "Myrtle Beach-Socastee, SC-NC", "Wilmington, NC",
          "Winston-Salem, NC")

df2<-df |>
  filter(add %in% cgroup) |>
  mutate(start=as.Date("01/11/2003", format="%d/%m/%Y")) |>
  mutate(open=as.Date("01/11/2007", format="%d/%m/%Y")) |>
  mutate(opentime=ifelse(as.Date(date)>open,1,0)) |>
  mutate(treatcity=ifelse(add=="Charlotte, NC",1,0)) |>
  mutate(time=interval(start, as.Date(date))/months(1)) |>
  mutate(add_num=factor(add, levels=cgroup)) |>
  mutate(add_num=as.numeric(add_num)) |>
  filter(time<=96) |>
  mutate(PRECTOT2=PRECTOT^2, PRECSNO2=PRECSNO^2, QSH2=QSH^2, RHOA2=RHOA^2, 
         SPEED2=SPEED^2, TLML2=TLML^2)

dataprep.out <-
  dataprep(
    df2,
    predictors            = c("PRECTOT", "PRECSNO", "QSH", "RHOA", "SPEED", 
                              "TLML","PRECTOT2", "PRECSNO2", "QSH2", "RHOA2",
                              "SPEED2", "TLML2"),
    dependent             = "pm25",
    unit.variable         = "add_num",
    time.variable         = "time",
    unit.names.variable   = "add",
    treatment.identifier  = 1,
    controls.identifier   = c(2:11),
    time.predictors.prior = c(-46:48),
    time.optimize.ssr     = c(-45:16), #before construction time
    time.plot             = c(-45:96)
  )

synth.out <- synth(dataprep.out)

print(synth.tables   <- synth.tab(
  dataprep.res = dataprep.out,
  synth.res    = synth.out)
)

path.plot(synth.res    = synth.out,
          dataprep.res = dataprep.out,
          Ylab         = c("Y"),
          Xlab         = c("Months"),
          Legend       = c("Treated City","Synthetic City"),
          Legend.position = c("topleft")
)

abline(v   = 48,
       lty = 2)

gaps.plot(synth.res    = synth.out,
          dataprep.res = dataprep.out,
          Ylab         = c("Gap"),
          Xlab         = c("Year"),
          Ylim         = c(-5, 5),
          Main         = ""
)

abline(v   = 48,
       lty = 2)
```

### Phoenix-Mesa

```{r}
#| echo: false
#| message: false
#| warning: false

cgroup<-c("Phoenix-Mesa, AZ", "Flagstaff, AZ", "Sierra Vista, AZ", "Tucson, AZ")

df2<-df |>
  filter(add %in% cgroup) |>
  mutate(start=as.Date("01/12/2004", format="%d/%m/%Y")) |>
  mutate(open=as.Date("01/12/2008", format="%d/%m/%Y")) |>
  mutate(opentime=ifelse(as.Date(date)>open,1,0)) |>
  mutate(treatcity=ifelse(add=="Phoenix-Mesa, AZ",1,0)) |>
  mutate(time=interval(start, as.Date(date))/months(1)) |>
  mutate(add_num=factor(add, levels=cgroup)) |>
  mutate(add_num=as.numeric(add_num)) |>
  filter(time<=96)|>
  mutate(PRECTOT2=PRECTOT^2, PRECSNO2=PRECSNO^2, QSH2=QSH^2, RHOA2=RHOA^2, 
         SPEED2=SPEED^2, TLML2=TLML^2)

dataprep.out <-
  dataprep(
    df2,
    predictors            = c("PRECTOT", "PRECSNO", "QSH", "RHOA", "SPEED", 
                              "TLML","PRECTOT2", "PRECSNO2", "QSH2", "RHOA2",
                              "SPEED2", "TLML2"),
    dependent             = "pm25",
    unit.variable         = "add_num",
    time.variable         = "time",
    unit.names.variable   = "add",
    treatment.identifier  = 1,
    controls.identifier   = c(2:4),
    time.predictors.prior = c(-59:48),
    time.optimize.ssr     = c(-59:7), #before construction time
    time.plot             = c(1:96)
  )

synth.out <- synth(dataprep.out)

print(synth.tables   <- synth.tab(
  dataprep.res = dataprep.out,
  synth.res    = synth.out)
)

path.plot(synth.res    = synth.out,
          dataprep.res = dataprep.out,
          Ylab         = c("Y"),
          Xlab         = c("Months"),
          Legend       = c("Treated City","Synthetic City"),
          Legend.position = c("topleft")
)

abline(v   = 48,
       lty = 2)

```

## Regressions

Regressions with Charlotte and its 13 control cities.

Column 1 has month and year fixed effects.

Column 2 has month, year, and city fixed effects.

Column 1 has month and year fixed effects.

For all regressions, we restrict data up to 4 years after light rail opening date.

```{r}
#| warning: false
#| echo: false

library("lfe")
library("stargazer")

p<-read.csv("pm25_allcities.csv")
m<-read.csv("merra_allcities.csv")

df<-merge(p, m, by=c("date", "add", "month", "year"))

cgroup<-c("Charlotte, NC", "Asheville, NC", "Charleston-North Charleston, SC", 
          "Columbia, SC", "Concord, NC", "Durham, NC", "Fayetteville, NC", 
          "Greenville, SC", "Myrtle Beach-Socastee, SC-NC", "Wilmington, NC",
          "Winston-Salem, NC")

df1<-df |>
  filter(add %in% cgroup) |>
  mutate(start=as.Date("01/11/2003", format="%d/%m/%Y")) |>
  mutate(open=as.Date("01/11/2007", format="%d/%m/%Y")) |>
  mutate(opentime=ifelse(as.Date(date)>open,1,0)) |>
  mutate(treatcity=ifelse(add=="Charlotte, NC",1,0)) |>
  mutate(time=interval(start, as.Date(date))/months(1)) |>
  mutate(add_num=factor(add, levels=cgroup)) |>
  mutate(add_num=as.numeric(add_num)) |>
  filter(time<=96) |>
  mutate(PRECTOT2=PRECTOT^2, PRECSNO2=PRECSNO^2, QSH2=QSH^2, RHOA2=RHOA^2, 
         SPEED2=SPEED^2, TLML2=TLML^2) |>
  mutate(constime=ifelse(date>=as.Date("01/03/2005", format="%d/%m/%Y") &
                                               date<=open, 1, 0)) |>
  mutate(group=1)

cm1<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year, 
          data=df1)

cm2<-felm(log(pm25) ~ opentime:treatcity + opentime + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            add + month + year, 
          data=df1)

cm3<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
            constime:treatcity + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year, 
          data=df1)

stargazer(cm1, cm2, cm3, type="text", dep.var.labels=c("ln(PM25)"))
```

Regressions with Charlotte and its 3 closely-matched cities from synthetic control.

```{r}
#| warning: false
#| echo: false
cgroup<-c("Charlotte, NC", "Concord, NC", "Winston-Salem, NC", "Columbia, SC")

df2<-df1 |>
  filter(add %in% cgroup)

cm1<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year, 
          data=df2)

cm2<-felm(log(pm25) ~ opentime:treatcity + opentime + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            add + month + year, 
          data=df2)

cm3<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
            constime:treatcity + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year, 
          data=df2)

stargazer(cm1, cm2, cm3, type="text", dep.var.labels=c("ln(PM25)"))
```

Regressions with Houston and its 13 control cities.

```{r}
#| warning: false
#| echo: false

cgroup<-c("Houston, TX", "Austin, TX", "Beaumont, TX", "Brownsville, TX",
          "College Station-Bryan, TX", "Corpus Christi, TX", 
          "Denton-Lewisville, TX", 
          "El Paso, TX-NM", "Laredo, TX", "Waco, TX",
          "Lubbock, TX", "Odessa, TX", "San Antonio, TX")

df2<-df |>
  filter(add %in% cgroup) |>
  mutate(start=as.Date("01/01/2000", format="%d/%m/%Y")) |>
  mutate(open=as.Date("01/01/2004", format="%d/%m/%Y")) |>
  mutate(opentime=ifelse(as.Date(date)>open,1,0)) |>
  mutate(treatcity=ifelse(add=="Houston, TX",1,0)) |>
  mutate(time=interval(start, as.Date(date))/months(1)) |>
  mutate(add_num=factor(add, levels=cgroup)) |>
  mutate(add_num=as.numeric(add_num)) |>
  filter(time<=96) |>
  mutate(PRECTOT2=PRECTOT^2, PRECSNO2=PRECSNO^2, QSH2=QSH^2, RHOA2=RHOA^2, 
         SPEED2=SPEED^2, TLML2=TLML^2)|>
  mutate(constime=ifelse(date>=as.Date("01/03/2001", format="%d/%m/%Y") &
                                               date<=open, 1, 0)) |>
  mutate(group=2)

#alltime periods
cm1<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year, 
          data=df2)

cm2<-felm(log(pm25) ~ opentime:treatcity + opentime + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            add + month + year, 
          data=df2)

cm3<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
            constime:treatcity + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year, 
          data=df2)

stargazer(cm1, cm2, cm3, type="text", dep.var.labels=c("ln(PM25)"))
```

Regressions with the Twin Cities and its 13 control cities.

```{r}
#| warning: false
#| echo: false

cgroup<-c("Minneapolis-St. Paul, MN", "Beloit, WI-IL", "Duluth, MN-WI", 
          "Janesville, WI",
          "Rochester, MN", "Wausau, WI")

df3<-df |>
  filter(add %in% cgroup) |>
  mutate(start=as.Date("01/06/2000", format="%d/%m/%Y")) |>
  mutate(open=as.Date("01/06/2004", format="%d/%m/%Y")) |>
  mutate(opentime=ifelse(as.Date(date)>open,1,0)) |>
  mutate(treatcity=ifelse(add=="Minneapolis-St. Paul, MN",1,0)) |>
  mutate(time=interval(start, as.Date(date))/months(1)) |>
  mutate(add_num=factor(add, levels=cgroup)) |>
  mutate(add_num=as.numeric(add_num)) |>
  filter(time<=96)|>
  mutate(PRECTOT2=PRECTOT^2, PRECSNO2=PRECSNO^2, QSH2=QSH^2, RHOA2=RHOA^2, 
         SPEED2=SPEED^2, TLML2=TLML^2)|>
  mutate(constime=ifelse(date>=as.Date("01/02/2001", format="%d/%m/%Y") &
                                               date<=open, 1, 0)) |>
  mutate(group=3)

#alltime periods
cm1<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year, 
          data=df3)

cm2<-felm(log(pm25) ~ opentime:treatcity + opentime + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            add + month + year, 
          data=df3)

cm3<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
            constime:treatcity + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year, 
          data=df3)

stargazer(cm1, cm2, cm3, type="text", dep.var.labels=c("ln(PM25)"))
```

Regressions with Phoenix-Mesa, AZ

```{r}
#| warning: false
#| echo: false

cgroup<-c("Phoenix-Mesa, AZ", "Flagstaff, AZ", "Sierra Vista, AZ", "Tucson, AZ")

df4<-df |>
  filter(add %in% cgroup) |>
  mutate(start=as.Date("01/12/2004", format="%d/%m/%Y")) |>
  mutate(open=as.Date("01/12/2008", format="%d/%m/%Y")) |>
  mutate(opentime=ifelse(as.Date(date)>open,1,0)) |>
  mutate(treatcity=ifelse(add=="Phoenix-Mesa, AZ",1,0)) |>
  mutate(time=interval(start, as.Date(date))/months(1)) |>
  mutate(add_num=factor(add, levels=cgroup)) |>
  mutate(add_num=as.numeric(add_num)) |>
  filter(time<=96)|>
  mutate(PRECTOT2=PRECTOT^2, PRECSNO2=PRECSNO^2, QSH2=QSH^2, RHOA2=RHOA^2, 
         SPEED2=SPEED^2, TLML2=TLML^2) |>
  mutate(constime=ifelse(date>=as.Date("01/07/2005", format="%d/%m/%Y") &
                                               date<=open, 1, 0)) |>
  mutate(group=4)

#alltime periods
cm1<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year, 
          data=df4)

cm2<-felm(log(pm25) ~ opentime:treatcity + opentime + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            add + month + year, 
          data=df4)

cm3<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
            constime:treatcity + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year, 
          data=df4)

stargazer(cm1, cm2, cm3, type="text", dep.var.labels=c("ln(PM25)"))
```

Regression with all treated and control cities.

Column 1 and 3 have month, year, and city group fixed effects. Standard errors are clustered at the group level.

Column 2 has month, year, and city fixed effects. Standard errors are clustered at the city level.

```{r}
#| warning: false
#| echo: false

dfall<-rbind(df1, df2, df3, df4)

cm1<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year + group | 0 | group,
          data=dfall)

cm2<-felm(log(pm25) ~ opentime:treatcity + opentime + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year + add | 0 | add,
          data=dfall)

cm3<-felm(log(pm25) ~ opentime:treatcity + opentime + treatcity +
            constime:treatcity + 
                 PRECTOT + PRECSNO + QSH + RHOA + SPEED + TLML +
                 PRECTOT2 + PRECSNO2 + QSH2 + RHOA2 + SPEED2 + TLML2 |
            month + year + group | 0 | group, 
          data=dfall)

stargazer(cm1, cm2, cm3, type="text", dep.var.labels=c("ln(PM25)"))
```

# References

Di, Q., Y. Wei, A. Shtein, C. Hultquist, X. Xing, H. Amini, L. Shi, I. Kloog, R. Silvern, J. Kelly, M. B. Sabath, C. Choirat, P. Koutrakis, A. Lyapustin, Y. Wang, L. J. Mickley & J. Schwartz. 2021. Daily and Annual PM2.5 Concentrations for the Contiguous United States, 1-km Grids, v1 (2000 - 2016). Palisades, New York: NASA Socioeconomic Data and Applications Center (SEDAC). <https://doi.org/10.7927/0rvr-4538>

Fageda, X. (2021). Do light rail systems reduce traffic externalities? Empirical evidence from mid-size European cities. Transportation Research Part D: Transport and Environment, 92, 102731.

Ham, W., Vijayan, A., Schulte, N., & Herner, J. D. (2017). Commuter exposure to PM2. 5, BC, and UFP in six common transport microenvironments in Sacramento, California. Atmospheric Environment, 167, 335-345.

Houston, D., Dang, A., Wu, J., Chowdhury, Z., & Edwards, R. (2016). The cost of convenience; air pollution and noise on freeway and arterial light rail station platforms in Los Angeles. Transportation Research Part D: Transport and Environment, 49, 127-137.

Mulley, C., Hensher, D. A., & Cosgrove, D. (2017). Is rail cleaner and greener than bus?. Transportation Research Part D: Transport and Environment, 51, 14-28.

Park, E. S., & Sener, I. N. (2019). Traffic-related air emissions in Houston: Effects of light-rail transit. Science of the Total Environment, 651, 154-161.

Shaver, K. (2013, October 23). Residents concerned about impact of proposed light-rail Purple Line. The Washington Post. [https://www.washingtonpost.com/](https://www.washingtonpost.com/local/trafficandcommuting/residents-concerned-about-impacts-of-proposed-light-rail-purple-line/2013/10/23/ffc25d52-3bea-11e3-a94f-b58017bfee6c_story.html)

Texas A&M Transportation Institute. (2016). Light-Rail Transit (LRT) – Transportation Policy Research. <https://policy.tti.tamu.edu/strategy/light-rail-transit/>
