---
title: "Treatment City Selection"
format: html
editor: visual
---

## Selection Criteria

-   As Di et al. (2021) provides data from 2000 to 2016, candidate treatment cities must have light rails that opened from January 2004 to December 2012 in order to have four years of data before and after light rail opening.

-   Treatment cities must have no increase heavy rail usage in the four years before and after opening window.

## Evaluation Window

| City          | Opening Month | Eval Start | Eval Ends |
|---------------|---------------|------------|-----------|
| Charlotte, NC | Nov 2007      | Nov 2004   | Nov 2011  |
| Houston, TX   | Jan 2004      | Jan 2001   | Jan 2008  |
| Minn, MN      | Jun 2004      | Jun 2001   | Jun 2008  |
| Phoenix, AZ   | Dec 2008      | Dec 2005   | Dec 2012  |

## 2013 Statistics

```{r}
library("tidyverse")

df <- read.csv("NTD/2013_Operating_Stats.csv")

bus_types<-c("MB", "RB", "CB", "TB")
oth_types<-c("CC", "FB", "IP", "TR", "DR")
rail_types<-c("CR", "HR", "YR", "MG", "SR")
car_types<-c("DT", "VP", "JT")

df2<-df |>
  select(ID, State, Urbanized.Area, Mode, Unlinked.Passenger.Trips) |>
  mutate(UPT=str_remove_all(Unlinked.Passenger.Trips, ",")) |>
  mutate(UPT=as.numeric(UPT)) |>
  mutate(mode_group=ifelse(Mode %in% bus_types, "bus", NA)) |>
  mutate(mode_group=ifelse(Mode %in% oth_types, "oth", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% rail_types, "rail", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% car_types, "cars", mode_group)) |>
  mutate(mode_group=ifelse(Mode=="LR", "LR", mode_group)) |>
  group_by(mode_group, ID, State, Urbanized.Area) |>
  summarize(UPT=sum(UPT))

df3<-pivot_wider(df2, 
                 names_from = "mode_group", 
                 values_from = "UPT",
                 values_fill=0) |>
  mutate(total=LR+bus+cars+oth+rail) |>
  mutate(share_light=LR/total, share_rail=rail/total, share_bus=bus/total)

df_y13<-df3 |>
  mutate(year=2013)

#NC, TX, MN
id<-c(4008, 6008, 5027)
```

By 2013, Phoenix, Charlotte, Houston still has no heavy rail. Minneapolis has 0.8% heavy rail use.

## 2004 Statistics

```{r}
df <- read.csv("NTD/2004_Operating_Stats.csv")

df2<-df |>
  select(ID, Mode, Unlinked.Passenger.Trips) |>
  mutate(UPT=str_remove_all(Unlinked.Passenger.Trips, ",")) |>
  mutate(UPT=as.numeric(UPT)) |>
  mutate(mode_group=ifelse(Mode %in% bus_types, "bus", NA)) |>
  mutate(mode_group=ifelse(Mode %in% oth_types, "oth", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% rail_types, "rail", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% car_types, "cars", mode_group)) |>
  mutate(mode_group=ifelse(Mode=="LR", "LR", mode_group)) |>
  group_by(mode_group, ID) |>
  summarize(UPT=sum(UPT))

df3<-pivot_wider(df2, 
                 names_from = "mode_group", 
                 values_from = "UPT",
                 values_fill=0) |>
  mutate(total=LR+bus+cars+oth+rail) |>
  mutate(share_light=LR/total, share_rail=rail/total, share_bus=bus/total)

df_y04<-df3 |>
  mutate(year=2004)
```

## 2005 Statistics

```{r}
df <- read.csv("NTD/2005_Operating_Stats.csv")

df2<-df |>
  select(ID, Mode, Unlinked.Passenger.Trips) |>
  mutate(UPT=str_remove_all(Unlinked.Passenger.Trips, ",")) |>
  mutate(UPT=as.numeric(UPT)) |>
  mutate(mode_group=ifelse(Mode %in% bus_types, "bus", NA)) |>
  mutate(mode_group=ifelse(Mode %in% oth_types, "oth", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% rail_types, "rail", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% car_types, "cars", mode_group)) |>
  mutate(mode_group=ifelse(Mode=="LR", "LR", mode_group)) |>
  group_by(mode_group, ID) |>
  summarize(UPT=sum(UPT))

df3<-pivot_wider(df2, 
                 names_from = "mode_group", 
                 values_from = "UPT",
                 values_fill=0) |>
  mutate(total=LR+bus+cars+oth+rail) |>
  mutate(share_light=LR/total, share_rail=rail/total, share_bus=bus/total)

df_y05<-df3 |>
  mutate(year=2005)
```

## 2006 Statistics

```{r}
df <- read.csv("NTD/2006_Operating_Stats.csv")

df2<-df |>
  select(ID, Mode, Unlinked.Passenger.Trips) |>
  mutate(UPT=str_remove_all(Unlinked.Passenger.Trips, ",")) |>
  mutate(UPT=as.numeric(UPT)) |>
  mutate(mode_group=ifelse(Mode %in% bus_types, "bus", NA)) |>
  mutate(mode_group=ifelse(Mode %in% oth_types, "oth", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% rail_types, "rail", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% car_types, "cars", mode_group)) |>
  mutate(mode_group=ifelse(Mode=="LR", "LR", mode_group)) |>
  group_by(mode_group, ID) |>
  summarize(UPT=sum(UPT))

df3<-pivot_wider(df2, 
                 names_from = "mode_group", 
                 values_from = "UPT",
                 values_fill=0) |>
  mutate(total=LR+bus+cars+oth+rail) |>
  mutate(share_light=LR/total, share_rail=rail/total, share_bus=bus/total)

df_y06<-df3 |>
  mutate(year=2006)
```

## 2007 Statistics

```{r}
df <- read.csv("NTD/2007_Operating_Stats.csv")

df2<-df |>
  select(ID, State, Name, Mode, Unlinked.Passenger.Trips) |>
  mutate(UPT=str_remove_all(Unlinked.Passenger.Trips, ",")) |>
  mutate(UPT=as.numeric(UPT)) |>
  mutate(mode_group=ifelse(Mode %in% bus_types, "bus", NA)) |>
  mutate(mode_group=ifelse(Mode %in% oth_types, "oth", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% rail_types, "rail", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% car_types, "cars", mode_group)) |>
  mutate(mode_group=ifelse(Mode=="LR", "LR", mode_group)) |>
  group_by(mode_group, ID, State, Name) |>
  summarize(UPT=sum(UPT))

df3<-pivot_wider(df2, 
                 names_from = "mode_group", 
                 values_from = "UPT",
                 values_fill=0) |>
  mutate(total=LR+bus+cars+oth+rail) |>
  mutate(share_light=LR/total, share_rail=rail/total, share_bus=bus/total)

df_y07<-df3 |>
  mutate(year=2007)
```

## 2008 Statistics

```{r}
df <- read.csv("NTD/2008_Operating_Stats.csv")

df2<-df |>
  select(ID, State, Name, Mode, Unlinked.Passenger.Trips) |>
  mutate(UPT=str_remove_all(Unlinked.Passenger.Trips, ",")) |>
  mutate(UPT=as.numeric(UPT)) |>
  mutate(mode_group=ifelse(Mode %in% bus_types, "bus", NA)) |>
  mutate(mode_group=ifelse(Mode %in% oth_types, "oth", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% rail_types, "rail", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% car_types, "cars", mode_group)) |>
  mutate(mode_group=ifelse(Mode=="LR", "LR", mode_group)) |>
  group_by(mode_group, ID, State, Name) |>
  summarize(UPT=sum(UPT))

df3<-pivot_wider(df2, 
                 names_from = "mode_group", 
                 values_from = "UPT",
                 values_fill=0) |>
  mutate(total=LR+bus+cars+oth+rail) |>
  mutate(share_light=LR/total, share_rail=rail/total, share_bus=bus/total)

df_az<-df3 |>
  filter(State=="AZ")

id_az<-c(9032, 9034, 9131, 9136, 9172, 9169, 9132, 9140, 9209)

df_y08<-df3 |>
  mutate(year=2008)

az_y08<-df3 |>
  filter(ID %in% id_az)|>
  mutate(year=2008)
```

## 2009 Statistics

```{r}
df <- read.csv("NTD/2009_Operating_Stats.csv")

df2<-df |>
  select(ID, State, Name, Mode, Unlinked.Passenger.Trips) |>
  mutate(UPT=str_remove_all(Unlinked.Passenger.Trips, ",")) |>
  mutate(UPT=as.numeric(UPT)) |>
  mutate(mode_group=ifelse(Mode %in% bus_types, "bus", NA)) |>
  mutate(mode_group=ifelse(Mode %in% oth_types, "oth", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% rail_types, "rail", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% car_types, "cars", mode_group)) |>
  mutate(mode_group=ifelse(Mode=="LR", "LR", mode_group)) |>
  group_by(mode_group, ID, State, Name) |>
  summarize(UPT=sum(UPT))

df3<-pivot_wider(df2, 
                 names_from = "mode_group", 
                 values_from = "UPT",
                 values_fill=0) |>
  mutate(total=LR+bus+cars+oth+rail) |>
  mutate(share_light=LR/total, share_rail=rail/total, share_bus=bus/total)

df_y09<-df3 |>
  mutate(year=2009)

az_y09<-df3 |>
  filter(ID %in% id_az)|>
  mutate(year=2009)
```

## 2010 Statistics

```{r}
df <- read.csv("NTD/2010_Operating_Stats.csv")

df2<-df |>
  select(ID, State, Name, Mode, Unlinked.Passenger.Trips) |>
  mutate(UPT=str_remove_all(Unlinked.Passenger.Trips, ",")) |>
  mutate(UPT=as.numeric(UPT)) |>
  mutate(mode_group=ifelse(Mode %in% bus_types, "bus", NA)) |>
  mutate(mode_group=ifelse(Mode %in% oth_types, "oth", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% rail_types, "rail", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% car_types, "cars", mode_group)) |>
  mutate(mode_group=ifelse(Mode=="LR", "LR", mode_group)) |>
  group_by(mode_group, ID, State, Name) |>
  summarize(UPT=sum(UPT))

df3<-pivot_wider(df2, 
                 names_from = "mode_group", 
                 values_from = "UPT",
                 values_fill=0) |>
  mutate(total=LR+bus+cars+oth+rail) |>
  mutate(share_light=LR/total, share_rail=rail/total, share_bus=bus/total)

df_y10<-df3 |>
  mutate(year=2010)

az_y10<-df3 |>
  filter(ID %in% id_az)|>
  mutate(year=2010)
```

## 2011 Statistics

```{r}
df <- read.csv("NTD/2011_Operating_Stats.csv")

df2<-df |>
  select(ID, State, Name, Mode, Unlinked.Passenger.Trips) |>
  mutate(UPT=str_remove_all(Unlinked.Passenger.Trips, ",")) |>
  mutate(UPT=as.numeric(UPT)) |>
  mutate(mode_group=ifelse(Mode %in% bus_types, "bus", NA)) |>
  mutate(mode_group=ifelse(Mode %in% oth_types, "oth", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% rail_types, "rail", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% car_types, "cars", mode_group)) |>
  mutate(mode_group=ifelse(Mode=="LR", "LR", mode_group)) |>
  group_by(mode_group, ID, State, Name) |>
  summarize(UPT=sum(UPT))

df3<-pivot_wider(df2, 
                 names_from = "mode_group", 
                 values_from = "UPT",
                 values_fill=0) |>
  mutate(total=LR+bus+cars+oth+rail) |>
  mutate(share_light=LR/total, share_rail=rail/total, share_bus=bus/total)

df_y11<-df3 |>
  mutate(year=2011)

az_y11<-df3 |>
  filter(ID %in% id_az)|>
  mutate(year=2011)
```

## 2012 Statistics

```{r}
df <- read.csv("NTD/2012_Operating_Stats.csv")

df2<-df |>
  select(ID, State, Name, Mode, Unlinked.Passenger.Trips) |>
  mutate(UPT=str_remove_all(Unlinked.Passenger.Trips, ",")) |>
  mutate(UPT=as.numeric(UPT)) |>
  mutate(mode_group=ifelse(Mode %in% bus_types, "bus", NA)) |>
  mutate(mode_group=ifelse(Mode %in% oth_types, "oth", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% rail_types, "rail", mode_group)) |>
  mutate(mode_group=ifelse(Mode %in% car_types, "cars", mode_group)) |>
  mutate(mode_group=ifelse(Mode=="LR", "LR", mode_group)) |>
  group_by(mode_group, ID, State, Name) |>
  summarize(UPT=sum(UPT))

df3<-pivot_wider(df2, 
                 names_from = "mode_group", 
                 values_from = "UPT",
                 values_fill=0) |>
  mutate(total=LR+bus+cars+oth+rail) |>
  mutate(share_light=LR/total, share_rail=rail/total, share_bus=bus/total)

df_y11<-df3 |>
  mutate(year=2012)

az_y12<-df3 |>
  filter(ID %in% id_az)|>
  mutate(year=2012)
```

## Charlotte, NC Light Rail Shares

```{r}
charlotte<-rbind(df_y07, df_y08, df_y09, df_y10, df_y11) |>
  filter(State=="NC")

ggplot(data=charlotte)+
  geom_line(aes(y=share_light, x=year, col="light rail"))+
  geom_line(aes(y=share_bus, x=year, col="bus"))
```

## Phoenix, AZ Light Rail Shares

```{r}
az<-rbind(az_y08, az_y09, az_y10, az_y11, az_y12) |>
  group_by(year) |>
  summarize(bus=sum(bus), 
            cars=sum(cars), 
            LR=sum(LR), 
            rail=sum(rail), 
            total=sum(total)) |>
  mutate(share_bus=bus/total, share_rail=rail/total, share_light=LR/total)

ggplot(data=az)+
  geom_line(aes(y=share_light, x=year, col="light rail"))+
  geom_line(aes(y=share_rail, x=year, col="rail"))+
  geom_line(aes(y=share_bus, x=year, col="bus"))
```

## MN Light Rail Shares

```{r}
df_y072<-df_y07 |>
  ungroup() |>
  dplyr::select(-State) |>
  dplyr::select(-Name)

mn<-rbind(df_y04, df_y05, df_y06, df_y072) |>
  filter(ID==5027)

ggplot(data=mn)+
  geom_line(aes(y=share_light, x=year, col="light rail"))+
  geom_line(aes(y=share_bus, x=year, col="bus"))
```

## Houston, TX Light Rail Shares

```{r}
tx<-rbind(df_y04, df_y05, df_y06, df_y072) |>
  filter(ID==6008)

ggplot(data=tx)+
  geom_line(aes(y=share_light, x=year, col="light rail"))+
  geom_line(aes(y=share_bus, x=year, col="bus"))
```

## Control cities construction

```{r}
df_y072<-df_y07 |>
  ungroup() |>
  dplyr::select(-State, -Name) 

df_y082<-df_y08 |>
  ungroup() |>
  dplyr::select(-State, -Name) 

df_y092<-df_y09 |>
  ungroup() |>
  dplyr::select(-State, -Name) 

df_y102<-df_y10 |>
  ungroup() |>
  dplyr::select(-State, -Name) 

df_y112<-df_y11 |>
  ungroup() |>
  dplyr::select(-State, -Name) 

cntrl<-rbind(df_y04, df_y05, df_y06, df_y072, df_y082, df_y092, df_y102, df_y112) |>
  group_by(ID, year) |>
  summarize(bus=sum(bus), 
            LR=sum(LR), 
            rail=sum(rail), 
            oth=sum(oth), 
            total=sum(total))
```
